Variable Selection in Convex Function Estimation

### 2013-10-04 tex/scam/appendix.tex
appendix

Fix $x_k$, we will show that the value $\E[ f(X) | x_k] - \mu^*$, for all $x_k$,  uniquely minimizes 
\[
\min_{ f_k(x_k) } \int_{\mathbf{x}_{-k}} p(\mathbf{x}) 
         \big| f(\mathbf{x}) - \sum_{k' \neq k} f^*_{k'} (x_{k'}) - f_k (x_k)\big|^2 d \mathbf{x}_{-k}.
\]
It easily follows then that the function $x_k \mapsto \E[ f(X) \given x_k] - \mu^*$ is the unique $f^*_k$ that minimizes the expected square error.

@@

we can re-write the Lagrangian $L$, in term of just $d_k,c_k$, as the following.
\[
\min_{d_k, c_k}  \frac{1}{2n} \| r_k - \Delta_k d_k + c_k \mathbf{1} \|_2^2 
        + \lambda \sum_{i=2}^n d_{ki} + \lambda|d_{k1}| 
        - \mu_k^\tran d_k + \gamma_k (c_k - \mathbf{1}_n^\tran\Delta_k d_k)
\]
where $r_k \coloneqq Y - \bar{Y}\mathbf{1}_n - \sum_{k' \in S, k' \neq k} (\Delta_{k'} d_{k'} - c_{k'} \mathbf{1}_n) $, and $\mu_k \in \R^n$ is a vector of dual variables where $\mu_{k,1} = 0$ and $\mu_{k,i} \geq 0$ for $i=2,...,n$.

@@

Suppose
\[
\lambda \mathbf{1} > |\frac{1}{n} \Delta_k^\tran\hat{r}|,
\]
then we easily see that the first coordinate of $\mathbf{u}_k$ can be set to some value in $(-1,1)$ and we can set $\mu_{k,i} > 0$ for $i=2,...,n$.

@@

\begin{align*}
\frac{1}{n} \hat{r}^\tran \Delta_{k,j} &= 
              \frac{1}{n} \hat{r}^\tran \max(X_k, X_{k(j)}\mathbf{1}_n) 
                                       - \frac{1}{n} \hat{r}^\tran \mathbf{1}_n X_{k(j)} \\
                  &= \frac{1}{n} \hat{r}^\tran \max(X_k, X_{k(j)})
\end{align*}
Our goal in this proof is to bound $\frac{1}{n} \hat{r}^\tran \max(X_k, X_{k(j)})$ from above.

@@

\begin{align*}
\hat{r}_i &= Y_i - \bar{Y}_i -  \sum_{k \in S} \hat{f}_k (X_k^{(i)}) \\
	&= f_0(X_S^{(i)}) + \epsilon_i - \bar{f_0} - \bar{\epsilon}- 
                      \sum_{k \in S} \hat{f}_k(X_k^{(i)}) \\
	&=  f_0(X_S^{(i)}) - \bar{f_0} - \sum_{k \in S} \hat{f}_k(X_k^{(i)}) 
                   + \epsilon_i - \bar{\epsilon}
\end{align*}
Where $\bar{f_0} = \frac{1}{n} \sum_{i=1}^n f_0(X_S^{(i)})$ and likewise for $\bar{\epsilon}$.

@@

 Given samples $X^{(1)},...,X^{(n)}$, let $f, g$ be a function and $w$ be a $n$-dimensional random vector, then we denote $\| f - g + w \|_n^2 \coloneqq \frac{1}{n} \sum_{i=1}^n ( f(X^{(i)}) - g(X^{(i)}) + w_i )^2$.

@@

We will argue that all the components of the additive function $f^{*(s-)}$ are also in $f^{*s}$. Let us denote the components of $f^{*s} = \sum_{k=1}^s f_k^*$. We will now invoke Lemma~\ref{lem:additive_component}, which is valid because we assume $X_1,...,X_s$ are independent by assumption A1'. By Lemma~\ref{lem:additive_component}, if we set $f^*_k = 0$, the resulting additive function $\sum_{k'\neq k} f_{k'}^*$ minimizes the population risk subject to the constraint that $f_k = 0$. By definition, $f^{*s}$ is $\argmin_k \sum_{k'\neq k} f_{k'}^*$ and thus share components with $f^{*s}$.

@@

\begin{align*}
|\hat{R}_s(f^{*s} - \bar{f}^{*s}) - \hat{R}_s(f^{*s})| &\leq 
               \| f_0 + w - \bar{f}_0 \mathbf{1}_n - \bar{w}\mathbf{1}_n 
                     - f^{*s} + \bar{f}^{*s}\mathbf{1}_n \|_n^2  - \| y - f^{*s}\|_n^2 \\
	&\leq 2\| f_0 + w - \bar{f}_0 \mathbf{1}_n - \bar{w} \mathbf{1}_n -f^{*s}\|_n 
                \|\bar{f}^{*s} \mathbf{1}_n \|_n 
                  + \|\bar{f}^{*s} \mathbf{1}_n \|_n^2 \\
        &\leq 2 | \bar{f}^{*s} | \| f_0 + w 
           - \bar{f}_0 \mathbf{1}_n - \bar{w} \mathbf{1}_n - f^{*s} \|_n + |\bar{f}^{*s}|^2 
\end{align*}

@@

\begin{align*}
\| f_0 + w -\bar{f}_0\mathbf{1}_n 
                       - \bar{w} \mathbf{1}_n - f^{*s} \|_n 
	\leq \| f_0 - f^{*s} \|_n + \|w\|_n + |\bar{f}_0| + |\bar{w}|\\
\end{align*}

@@

For a fixed $f'$, we have that, by definition
\begin{align*}
\hat{R}_s (f') &= \| f_0 + w - \bar{f}_0\mathbf{1}_n - \bar{w} \mathbf{1}_n - f' \|_n^2\\
     &= \| f_0  - \bar{f}_0\mathbf{1}_n - \bar{w}\mathbf{1}_n - f'\|_n^2 
       + 2 \langle w, f_0 - \bar{f}_0\mathbf{1}_n -\bar{w}\mathbf{1}_n - f'\rangle_n 
       + \| w \|_n^2 \\
R_s(f') &= \E\big(f_0(X) + w - \mu - f'(X)\big)^2 \\
     &= \E\big(f_0(X) - \mu - f'(X)\big)^2 + \E w^2  
\end{align*}

Because $f_0(X^{(i)}) - f'^(X^{(i)})$ is bounded by $3sB$, $\| f_0 - f'\|_n^2$ is the empirical average of $n$ random variables bounded by $9(sB)^2$.

@@

Collecting all these results and applying union bound, we have that $\sup_{f' \in \mathcal{C}_\epsilon[s,b,B,L]} | \hat{R}_s(f') - R_s(f') | \geq t$ occurs with probability at most
\[
C \exp\big( s\left( \frac{bBLs}{\epsilon} \right)^{1/2} - c n t^2 \frac{1}{\sigma^2 (sB)^4} \big)
\]
for all $n > n_0$.

@@

Let $f' \coloneqq = \sum_{k=1}^s f'_k$, then 
\[
\| f' - f \|_{\infty} \leq \sum_{k=1}^s \| f_k - f'_k \|_\infty \leq s \frac{\epsilon}{s} \leq \epsilon
\]

>>>>

Fix $x_k$, we will show that the value $\E[ f(X) | x_k] - \mu^*$, for all $x_k$,  uniquely minimizes 
\[
\min_{ f_k(x_k) } \int_{\mathbf{x}_{-k}} p(\mathbf{x}) 
         \big| f(\mathbf{x}) - \sum_{k' \neq k} f^*_{k'} (x_{k'}) - f_k (x_k) -\mu^*\big|^2 
                 d \mathbf{x}_{-k}.
\]
It easily follows then that the function $x_k \mapsto \E[ f(X) \given x_k] - \mu^*$ is the unique $f^*_k$ that minimizes the expected square error.

@@

we can re-write the Lagrangian $L$, in term of just $d_k,c_k$, as the following.
\[
\min_{d_k, c_k}  \frac{1}{2n} \| r_k - \Delta_k d_k + c_k \mathbf{1} \|_2^2 
        + \lambda \sum_{i=2}^n d_{ki} + \lambda|d_{k1}| 
        - \mu_k^\tran d_k + \gamma_k (c_k - \mathbf{1}_n^\tran\Delta_k d_k)
\]
where $r_k \coloneqq Y - \bar{Y}\mathbf{1}_n - \sum_{k' \in S, k' \neq k} (\Delta_{k'} d_{k'} - c_{k'} \mathbf{1}_n) $, and $\mu_k \in \R^{n-1}$ is a vector of dual variables where $\mu_{k,1} = 0$ and $\mu_{k,i} \geq 0$ for $i=2,...,n-1$.

@@

Suppose
\[
\lambda \mathbf{1} > |\frac{1}{n} \Delta_k^\tran\hat{r}|,
\]
then we easily see that the first coordinate of $\mathbf{u}_k$ can be set to some value in $(-1,1)$ and we can set $\mu_{k,i} > 0$ for $i=2,...,n-1$.

@@

\begin{align*}
\frac{1}{n} \hat{r}^\tran \Delta_{k,j} &= 
              \frac{1}{n} \hat{r}^\tran \max(X_k, X_{k(j)}\mathbf{1}_n) 
                                       - \frac{1}{n} \hat{r}^\tran \mathbf{1}_n X_{k(j)} \\
                  &= \frac{1}{n} \hat{r}^\tran \max(X_k, X_{k(j)}\mathbf{1}_n)
\end{align*}
Our goal in this proof is to bound $\frac{1}{n} \hat{r}^\tran \max(X_k, X_{k(j)})$ from above.

@@

\begin{align*}
\hat{r}_i &= Y_i - \bar{Y} -  \sum_{k \in S} \hat{f}_k (X_k^{(i)}) \\
	&= f_0(X_S^{(i)}) + \epsilon_i - \bar{f_0} - \bar{\epsilon}- 
                      \sum_{k \in S} \hat{f}_k(X_k^{(i)}) \\
	&=  f_0(X_S^{(i)}) - \bar{f_0} - \sum_{k \in S} \hat{f}_k(X_k^{(i)}) 
                   + \epsilon_i - \bar{\epsilon}
\end{align*}
Where $\bar{f_0} = \frac{1}{n} \sum_{i=1}^n f_0(X_S^{(i)})$ and likewise for $\bar{\epsilon}$.

@@

 Given samples $X^{(1)},...,X^{(n)}$, let $f, g$ be a function and $w$ be a $n$-dimensional random vector, then we denote $\| f - g + w\|_n^2 \coloneqq \frac{1}{n} \sum_{i=1}^n ( f(X^{(i)}) - g(X^{(i)}) + w_i )^2$. We will also abuse notation and let $\| f + c \|_n^2 = \frac{1}{n} \sum_{i=1}^n (f(X^{(i)}) + c)^2$ if $c$ is a scalar.

We let $\langle f, g \rangle_n \coloneqq \frac{1}{n} \sum_{i=1}^n f(X^{(i)}) g(X^{(i)})$. It then follows that:
\begin{packed_enum}
\item $\| f + g \|_n^2 = \|f\|_n^2 + 2 \langle f, g \rangle_n + \|g \|_n^2$
\item $\langle f, g \rangle_n \leq \|f\|_n \| g \_n$
\end{packed_enum}

@@

We will argue that all the components of the additive function $f^{*(s-1)}$ are also in $f^{*s}$. Let us denote the components of $f^{*s} = \sum_{k=1}^s f_k^*$. We will now invoke Lemma~\ref{lem:additive_component}, which is valid because we assume $X_1,...,X_s$ are independent by assumption A1'. By Lemma~\ref{lem:additive_component}, if we set $f^*_k = 0$, the resulting additive function $\sum_{k'\neq k} f_{k'}^*$ minimizes the population risk subject to the constraint that $f_k = 0$. By definition, $f^{*s}$ is $\argmin_k \sum_{k'\neq k} f_{k'}^*$ and thus share components with $f^{*s}$.

@@

\begin{align*}
|\hat{R}_s(f^{*s} - \bar{f}^{*s}) - \hat{R}_s(f^{*s})| &\leq 
         \| f_0 + w - \bar{f}_0 - \bar{w}
         - f^{*s} + \bar{f}^{*s} \|_n^2  - \| f_0 + w - \bar{f}_0 - \bar{w} - f^{*s}\|_n^2 \\
   &\leq 2 \langle f_0 + w - \bar{f}_0 - \bar{w} - f^{*s}, \bar{f}^{*s} \rangle_n 
         + \| \bar{f}^{*s} \|_n^2 \\
   &\leq 2\| f_0 + w - \bar{f}_0 - \bar{w} -f^{*s}\|_n 
                \|\bar{f}^{*s} \|_n 
                  + \|\bar{f}^{*s} \|_n^2 \\
   &\leq 2 | \bar{f}^{*s} | \| f_0 + w 
           - \bar{f}_0 - \bar{w} - f^{*s} \|_n + |\bar{f}^{*s}|^2 
\end{align*}

@@

\begin{align*}
\| f_0 + w -\bar{f}_0 - \bar{w} - f^{*s} \|_n 
	\leq \| f_0 - f^{*s} \|_n + \|w\|_n + |\bar{f}_0| + |\bar{w}|\\
\end{align*}

@@

For a fixed $f'$, we have that, by definition
\begin{align*}
\hat{R}_s (f') &= \| f_0 + w - \bar{f}_0 - \bar{w} - f' \|_n^2\\
      &= \|f_0 - \bar{f}_0 - f'\|_n^2 + 2 \langle w - \bar{w}, f_0 - \bar{f}_0 - f'\rangle_n
             + \| w - \bar{w}\|_n^2\\
R_s(f') &= \E\big(f_0(X) + w - \mu - f'(X)\big)^2 \\
     &= \E\big(f_0(X) - \mu - f'(X)\big)^2 + \E w^2 
\end{align*}

Therefore:
\begin{align*}
\hat{R}_s(f') - R_s(f') = & \|f_0 -\bar{f}_0 - f'\|_n^2 - \E(f_0(X) - \mu - f'(X))^2 \\
            &+ \|w - \bar{w}\|_n^2 - \E w^2 \\
            &+ 2\langle w-\bar{w}, f_0 - \bar{f}_0 - f'\rangle_n
\end{align*}

\textbf{Step 3.1:} We first bound $2 \langle w - \bar{w}, f_0 - \bar{f}_0 - f'\rangle_n$.
\begin{align*}
\langle w - \bar{w}, f_0 - \bar{f}_0 - f'\rangle_n &=
      \langle w, f_0 - \bar{f}_0 - f'\rangle_n 
      - \langle \bar{w}, f_0 - \bar{f}_0 - f'\rangle_n \\
\end{align*}

The first term, fully expanded, is $\frac{1}{n} \sum_{i=1}^n w_i (f_0(X^{(i)}) - \bar{f}_0 - f'(X^{(i)}))$. Since $w_i$ and $X^{(i)}$ are independent, we use the sub-Gaussian concentration inequality. Note that $|f_0(X^{(i)}) - \bar{f}_0 - f'(X^{(i)})| \leq 3sB$, and so $|\langle w, f_0 - \bar{f}_0 - f' \rangle| > t$ with probability at most $C \exp\big( - c n t^2 \frac{1}{\sigma^2 (sB)^2} \big)$.

The second term, fully expanded, is $\bar{w} \bar{f'}$. $\bar{f'} \leq sB$ and so $|\bar{w}\bar{f'}| > t$ with probability at most $C \exp\big( - c n t^2 \frac{1}{\sigma^2(sB)^2} \big)$ as well.

\textbf{Step 3.2} We now bound $\| w - \bar{w} \|_n^2 - \E w^2$. 
\begin{align*}
\|w - \bar{w} \|_n^2 &= \|w\|_n^2 - 2 \langle w, \bar{w} \rangle_n + \|\bar{w}\|_n \\
               &= \|w\|_n^2 - \bar{w}^2
\end{align*}

Using sub-Exponential concentration, we know that $| \|w\|_n^2 - \E w^2 | \geq t$ with probability at most $C \exp( - c n \frac{1}{\sigma^2})$.

$\bar{w} \leq \sigma\frac{1}{cn}$ with probability at least $1-\frac{C}{n}$.

\textbf{Step 3.3:} We now bound $\| f_0 - \bar{f}_0 - f'\|_n^2 - \E(f_0(X) - \mu - f'(X))^2$.

Because $f_0(X^{(i)}) - f'^(X^{(i)})$ is bounded by $2sB$, $\| f_0 - f'\|_n^2$ is the empirical average of $n$ random variables bounded by $9(sB)^2$.

@@

and applying union bound, we have that $\sup_{f' \in \mathcal{C}_\epsilon[s,b,B,L]} | \hat{R}_s(f') - R_s(f') | \geq t$ occurs with probability at most
\[
C \exp\big( s\left( \frac{bBLs}{\epsilon} \right)^{1/2} - c n t^2 \frac{1}{\sigma^2 (sB)^4} \big)
\]
for all $n > n_0$.

@@

Let $f' \coloneqq \sum_{k=1}^s f'_k$, then 
\[
\| f' - f \|_{\infty} \leq \sum_{k=1}^s \| f_k - f'_k \|_\infty \leq s \frac{\epsilon}{s} \leq \epsilon
\]
